---
title: "DATA 607 Project 1"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output: html_document
---
# Introduction
The overall nature of this project is to wrangle chess tournament data from a poorly formatted table. The data, which is loaded from a text file without standardized delimiters, requires the use of regular expression for extraction. A column of computed average opponent rating is also added to the data set. The tidyverse is used extensively throughout in stringr, dplyr and as well as occasional piping through magrittr. 

# Importing Packages

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(kableExtra)
```

# Loading File

Retrieving the raw data from Github for reproducibility.
```{r }
txt_data <- read_file("https://raw.githubusercontent.com/adavidowitz100/DATA607/main/Project1/tournamentinfo.txt")
print(substr(txt_data, 1, 1000))
```

# Data Discovery and Structuring

Using Regex in Stringr to extract the required data and adding to tidy dataframes. The matching text for the regex patern is commented in-line.

```{r}
player_name <- unlist(stringr::str_extract_all(txt_data,"[|]\\s+[A-Z]+\\s+[A-Z]+([A-Z]|\\s|[-])*\\s+[|]")) #| GARY HUA                      |
player_name <- unlist(stringr::str_extract_all(player_name,"[A-Z]+\\s+[A-Z]+([A-Z]|\\s|[-])*[A-Z]+"))             #GARY HUA
player_name <- as.data.frame(player_name)
state <- unlist(stringr::str_extract_all(txt_data,"[|]\\n\\s+[A-Z]{2}\\s+[|]")) #|\n   ON |
state <- unlist(stringr::str_extract_all(state,"[A-Z]{2}")) #ON
state <- as.data.frame(state)
point_total <- as.numeric(unlist(stringr::str_extract_all(txt_data,"\\d[.]\\d"))) #6.0
point_total <- as.data.frame(point_total)
pre_rating <- unlist(stringr::str_extract_all(txt_data,"R[:]\\s+\\d{3,4}")) #R: 1794
pre_rating <- as.integer(unlist(stringr::str_extract_all(pre_rating,"\\d{3,4}"))) #1794
pre_rating <- as.data.frame(pre_rating)
opponents <- unlist(stringr::str_extract_all(txt_data,"\\d[.]\\d.*[|]\\n\\s+[A-Z]{2}\\s+[|]")) #|W  39|W  21|W  18|W  14|W   7|D  12|D   4|
opponents <- unlist(stringr::str_remove(opponents, "\\d[.]\\d"))
opponents <- stringr::str_split(opponents,"[|]")
for(i in 1:length(opponents)){
opponents[[i]] <- as.integer(stringr::str_extract_all(opponents[[i]], "\\d{1,2}"))
}
opponents_df <- as.data.frame(do.call(rbind, opponents))
opponents_df <- opponents_df[-c(1,9,10)]

chess_df <- cbind.data.frame(player_name, state, point_total, pre_rating, opponents_df)
names(chess_df) <- c("Name", "State", "Points", "Pre-Rating", "Opponent 1","Opponent 2","Opponent 3","Opponent 4","Opponent 5","Opponent 6","Opponent 7")

chess_df %>% 
  head() %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling()

```

# Data Enriching

Computing the average opponents rating, adding it to the dataframe and then sub-setting for the final output.

```{r}
avg_vector <- character(nrow(player_name))
for(i in 1:nrow(player_name)){
  counter = 0
  total = 0
  for(j in 5:11){
    if(!is.na(chess_df[i,j])){
    counter <- counter + 1
    total <- total + chess_df[chess_df[i,j],"Pre-Rating"]
    }
  }
  avg_vector[i] <- round(total/counter, digits = 0)
}

chess_df <- cbind(chess_df, avg_vector)



csv_df <- dplyr::select(chess_df, Name, State, Points, "Pre-Rating", avg_vector)
csv_df <- rename(csv_df, "Avg Opponent Rating" = "avg_vector")

csv_df %>% 
  head() %>%
  kableExtra::kbl() %>%
  kableExtra::kable_styling()
```


# Output
Save the CSV to the working directory.

```{r}
directory = stringr::str_c(getwd(),"/Project1_Output.csv")
write.csv(csv_df,directory, row.names = FALSE)
```

# Conclusion
Given that the data is now in a widely accessible CSV format, this data can be used easily in further analysis. Perhaps, the overall average opponent ranking can be used to verify the pre-tournament ratings for players. While this data is perfectly usable as is, I would have liked to refactor some of the above code and specifically to include more piping in the stringr regex section.

