---
title: "DATA607 - Assignment 1"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
---

## Overview

Five Thirty Eight maintains power ratings for every MLB baseball team over all of MLB history. I subset this data to show the power ratings of only the NL East teams as the 2022 season has unfolded.

The current predictions are available at: <https://projects.fivethirtyeight.com/2022-mlb-predictions/>

Basic overview of data: <https://github.com/fivethirtyeight/data/tree/master/mlb-elo>

More in depth info on rating system can be found at: <https://fivethirtyeight.com/features/how-our-mlb-predictions-work/>

## Import libraries

```{r Import library, message=FALSE}
library(dplyr)
library(ggplot2)
```

## Read in the CSV of data from Github

```{r Reading CSV}
theUrl <- "https://raw.githubusercontent.com/adavidowitz100/DATA607/main/Assignment%201/mlb-elo/mlb_elo_latest.csv"
mlb_data <- read.table(file=theUrl, header=TRUE, sep=",")
```

## Clean and Subset Using dplyr

```{r clean and subset}
#Date needs to be formatted correctly
mlb_data$date <- as.Date(mlb_data$date)
#Subset teams to NL East
mlb_data <- select(mlb_data, date, team1, team2, rating1_pre, rating2_pre)
#convert data frame of games into data frame of teams
team1_data <- filter(mlb_data, team1 %in% c('NYM', 'ATL', 'PHI', 'FLA', 'WSN'))
team1_data <- select(team1_data, date, team1, rating1_pre)
names(team1_data)[names(team1_data)=="team1"] <- "team"
names(team1_data)[names(team1_data)=="rating1_pre"] <- "rating"
team2_data <-  filter(mlb_data,team2 %in% c('NYM', 'ATL', 'PHI', 'FLA', 'WSN'))
team2_data <- select(team2_data, date, team2, rating2_pre)
names(team2_data)[names(team2_data)=="team2"] <- "team"
names(team2_data)[names(team2_data)=="rating2_pre"] <- "rating"
NL_east <- rbind(team1_data, team2_data)
```

## Plot

```{r NL East Ratings Plot}
NL_east_plot <- ggplot(NL_east, aes(x=date, y=rating))
NL_east_plot <- NL_east_plot + geom_line(aes(color=factor(team), group=team))
NL_east_plot <- NL_east_plot + scale_color_discrete(name="Team")
NL_east_plot <- NL_east_plot + labs(title="2022 Power Rating of NL East Teams", x="Date", y="Rating")
NL_east_plot
```
