---
title: "Israeli Vaccination Data Analysis"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


## Introduction
Another example of the Simpson's paradox is apparent from analyzing the data from Israeli Covid-19 Vaccinations. When examining case counts it seems puzzling that the fully vaccinated have more cases in the over 50 population. Factoring rates per 100k of infections show the true efficacy of the vaccines.

## Import Libraries

```{r}
library(tidyverse)
```

## Import Data
Import the data from a CSV file stored on Github for reproducibility
```{r}
raw_csv <- readr::read_csv("https://raw.githubusercontent.com/adavidowitz100/DATA607/main/Israeli_Vaccinations_Extra_Credit/israeli_vaccination_data_analysis_start.csv")
print(raw_csv)
```

## Data Munging
Change the data from a wide contingency style table to a long tidy format
```{r}
vaxdf <- raw_csv[rowSums(is.na(raw_csv) | raw_csv == "") != ncol(raw_csv),] #remove blank rows
vaxdf <- vaxdf |> dplyr::filter(!is.na(Age)) |> #remove sub-headers and percentage rows
                  dplyr::rename("age" = 1, "unvax_pop" = 2, "vax_pop" = 3, "unvax_count" = 4, "vax_count" = 5, "efficacy" = 6) |> #rename columns
                  tidyr::pivot_longer(c("unvax_pop","vax_pop"),names_to = "status", values_to = "population") |> #combine pop
                  tidyr::pivot_longer(c("unvax_count","vax_count"),names_to = "status_count", values_to = "case_count") |> #combine case counts
                  dplyr::filter(stringr::str_sub(status,1,3) == stringr::str_sub(status_count,1,3)) |> #remove non-matching pop/case rows
                  dplyr::select(!status_count & !efficacy) #remove extra status column
vaxdf$status <- sub("\\_pop", "", vaxdf$status)
vaxdf$population <- as.integer(stringr::str_remove_all(vaxdf$population,",")) #remove , from pop and retype to int
vaxdf$case_count <- as.integer(vaxdf$case_count) #retype to int

vaxdf <- mutate(vaxdf, case_per_100k=case_count*100000/population) #add per 100k rate
```

## Analysis
```{r}
#calculate efficacy by age
vax_eff_df <- distinct(select(vaxdf,age)) |> left_join(dplyr::select(dplyr::filter(vaxdf,status == "vax"),age,case_per_100k), by="age") |> 
            rename(vax_case_rate = case_per_100k) |> 
            left_join(dplyr::select(dplyr::filter(vaxdf,status == "unvax"),age,case_per_100k), by="age") |> 
            rename(unvax_case_rate = case_per_100k) |>
            mutate(efficacy = 1-(vax_case_rate/unvax_case_rate))
                                           
#calculate efficacy for all
vax_eff_all_df <- vaxdf |> dplyr::select(status,population,case_count) |> dplyr::group_by(status) |>    dplyr::summarise(population=sum(population),case_count=sum(case_count))
vax_eff_all_df <- mutate(vax_eff_all_df, case_per_100k=case_count*100000/population) #add per 100k rate for all ages
vaxdf
vax_eff_df
total_eff <- as.numeric(1 -((dplyr::select(dplyr::filter(vax_eff_all_df,status == "vax"),case_per_100k))/(dplyr::select(dplyr::filter(vax_eff_all_df,status == "unvax"),case_per_100k))))
total_eff
```


## Conclusion
As you can see from the data the overall efficacy was high in each age group. The overall efficacy of 67% which is lower than the stratified efficacy is explained by the confounders introduced by age. It appears that Simpson's paradox is induced because the over 50 population had both higher vaccination rates as well as more overall serious poor health. In both cases these are expected and does not diminish the fact that each demographic was well protected by the vaccine shown by the considerably smaller rates of serious illness compared to the same age groups unvaccinated rate of serious illness.