---
title: "Assignment 4"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


## Introduction
This assignment loads, munges and analyses airline flight data. The data is initially messy and tools from the Tidyverse are used to standardize it into a "tidy" format. A plot is then used to highlight trends in the data.

## Import Libraries

```{r}
library(tidyverse)
```

## Load Data
Load the data from a CSV file stored on Github for reproducibility.
```{r}
theUrl <- "https://raw.githubusercontent.com/adavidowitz100/DATA607/main/Assignment4/Flight_Data.csv"
flight_data <- read.table(file=theUrl, header=TRUE, sep=",")
as_tibble(flight_data)
```

## Data Munging into Tidy Data

Remove blank rows
```{r}
flight_data <- flight_data[rowSums(is.na(flight_data) | flight_data == "") != ncol(flight_data),]
head(flight_data)
```
Rename Columns
```{r}
flight_data <- flight_data |> dplyr::rename("Airline" = X, "Status" = X.1)
head(flight_data)
```
Fill missing Airline values
```{r}
flight_data[flight_data==""]<-NA
flight_data <- tidyr::fill(flight_data, Airline, .direction = "down")
head(flight_data)
```
Convert data from wide format to Tidy long format by moving city counts from multiple columns to 2 new city and count name/value columns
```{r}
long_flight_data <- tidyr::pivot_longer(flight_data, c(Los.Angeles, Phoenix, San.Diego, San.Francisco, Seattle), names_to = "City", values_to = "Count")
head(long_flight_data)
```

Remove periods from City names
```{r}
long_flight_data$City <- sub("\\.", " ", long_flight_data$City)
head(long_flight_data)
```


## Analysis
Adding in a column for the percentage of flights with delays for each airline/city observation. This enables a better comparison of performance across the two airlines. A bar plot of the relative delays shows a clear trend.
```{r}
long_flight_data_delay <- long_flight_data |> dplyr::group_by(Airline, City) |> 
  dplyr::mutate(Delay_Percent = Count / sum(Count) * 100) |> 
  dplyr::filter(Status=="delayed") |> 
  dplyr::select(Airline, City, Delay_Percent)

ggplot(long_flight_data_delay, aes(x=City, y=Delay_Percent, fill=Airline)) +
    geom_bar(stat='identity', position='dodge')
```

## Conclusion
From examining the plotted ratios of flight delays, AM West has a significantly worse delay performance for the five cities in question. AM West on-time performance lagged behind Alaska Airlines in every city but had a significant gap in San Francisco and Seattle specifically. The analysis is hampered somewhat in scope because the data is an aggregated categorical data point of delayed/on-time. A numerical measure of delay times could yield more interesting results.  