---
title: "Assignment3"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(stringr)
```


## Question 1

Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

```{r Reading CSV, warning=FALSE, message=FALSE}
theUrl <- "https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv"
major_data <- read.table(file=theUrl, header=TRUE, sep=",")
ds_major <- major_data %>% dplyr::filter(stringr::str_detect(Major, "DATA|STATISTICS"))
print(ds_major)
```

## Question 2

Write code that transforms the data below:
[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"
Into a format like this:
c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

```{r}
unformatted_fruit <- '[1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"
[5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  
[9] "elderberry"   "lime"         "lychee"       "mulberry"    
[13] "olive"        "salal berry"'

expected_output <- 'c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")'

counter <- (stringr::str_count(unformatted_fruit,'\"[a-z]+\\s?[a-z]+\"'))
fruit <- stringr::str_extract_all(unformatted_fruit,'\"[a-z]+\\s?[a-z]+\"')
fruit_output <- "c("
for(i in 1:counter){
    fruit_output <- paste0(fruit_output, as.character(fruit[[1]][i]))
    if(i < counter){fruit_output <- paste0(fruit_output, ", ") }
}
fruit_output <- paste0(fruit_output, ")" )

print(fruit_output == expected_output)

cat(fruit_output)
```

## Question 3

Describe, in words, what these expressions will match:

    (.)\1\1 - Not a valid regex expression because you need another backslash in "\\1" to escape the string
    "(.)(.)\\2\\1" - A first letter, a second, second, first
    (..)\1 - Not a valid regex expression because you need another backslash in "\\1" to escape the string
    "(.).\\1.\\1" - A first letter, any letter, first letter, any letter, first letter
    "(.)(.)(.).*\\3\\2\\1" first letter, second, third, any number of letters, third, second, first


## Question 4 

Construct regular expressions to match words that:

    Start and end with the same character.
    "^(.).*\\1$"
    Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
    "(..).*\\1"
    Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)
    "(.).*\\1.*\\1"
    
```{r}
print(stringr::str_detect("designed", "^(.).*\\1$"))
print(stringr::str_detect("church", "(..).*\\1"))
print(stringr::str_detect("eleven", "(.).*\\1.*\\1"))
```
    

