---
title: "Project 4 - Part 1"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Introduction
My classmate Sanielle Worrell posted an excelent example of wide data generated by Zillow and proposed using Zillow data to analyze the change in New York City home prices over the last year. The data utilizes a Zillow methodology called ZHVI (Zillow Home Value Index) which represents "smoothed, seasonally adjusted measure of the typical home value and market changes across a given region and housing type... for homes in the 35th to 65th percentile range." Sanielle proposed doing the analysis with ZHVI by zipcode. However, my analysis below uses ZHVI by county due to the sheer quantity of data invloved (a 50MB CSV for zipcodes compared to 5MB for county). The methodology does not differ in any significant way from this change. The data is freely available at https://www.zillow.com/research/data/

## Import Libraries

```{r}
library(tidyverse)
```


## Load Data
Load the data from a CSV file stored on Github for reproducibility.
```{r}
theUrl <- "https://raw.githubusercontent.com/adavidowitz100/DATA607/main/Assignment4/Flight_Data.csv"
flight_data <- read.table(file=theUrl, header=TRUE, sep=",")
head(flight_data)


