---
title: "Project 2 - Part 2"
author: "Avery Davidowitz"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Introduction
This analysis of Pro Football Reference NFL standing data was propose by my classmate Andrew Bowen. The data is in tabular form but has sub-headers indicating the division of the below 4 teams. This format is not "Tidy" because each row should be an observation and this format uses a row to store a variable for multiple rows instead of in a dedicated column. This data set is somewhat contrived because Pro Football Reference does have an export CSV option on their site which removes the division data from the table. However, many similar data sources have tables with sub-headers like the NFL standings. I will then use passing yardage data to see if there is a correlation between passing yards and win percentage. The raw data is also available at https://www.pro-football-reference.com/years/2022/#AFC

## Import Libraries

```{r}
library(tidyverse)
```


## Load Data
Load the data from a CSV file stored on Github for reproducibility.
```{r}
theUrl <- "https://raw.githubusercontent.com/adavidowitz100/DATA607/main/Project2/football_reference_AFC.csv"
afc_data <- read.table(file=theUrl, header=TRUE, sep=",", na.strings = "", stringsAsFactors = FALSE)
head(afc_data)
```


## Data Munging into Tidy Data

This data can be easily tidied by adding a column for division and then filling down the values to each subsequent row. I can then add another column for Conference to split the two categorical variables of Conference and Division.

```{r}
afc_data <- afc_data |> dplyr::mutate(Division = ifelse(is.na(W),Tm, NA), .after = Tm) |>
  tidyr::fill(Division, .direction = "down") |>
  dplyr::mutate(Conference = "AFC", .after = Tm) |>
  dplyr::mutate(Division = stringr::str_sub(Division, start =  5)) |>
  dplyr::filter(!is.na(W))

head(afc_data)
```

## Loading Passing Data for Analysis

```{r}
theUrl <- "https://raw.githubusercontent.com/adavidowitz100/DATA607/main/Project2/football_reference_passing.csv"
passing_data <- read.table(file=theUrl, header=TRUE, sep=",", na.strings = "", stringsAsFactors = FALSE)
tail(passing_data)
```

## Data Munging into Tidy Data
This passing data is tidy except for three rows of summary data which needs to be removed.

```{r}
passing_data <- passing_data |> dplyr::filter(!is.na(Rk))
tail(passing_data)
```

## Joining Data
Joining win percentage and passing yards from both data sources.
```{r}
afc_pass_win <- afc_data |> dplyr::left_join(passing_data, by = "Tm") |>
  dplyr::select(W.L., Yds)
head(afc_pass_win)
```

## Plot
A plot will quickly show if there is a trend between passing yards and team win percentage.

```{r}
ggplot(afc_pass_win, aes(x=Yds, y=W.L.)) + geom_point()
```

## Conclusion
At this point in the season there does not seem to be a strong relationship between passing yards and win percentage. There are many teams with the same .500 win percentage on all ends of the passing yardage spectrum. However, there does seem to be a slight relationship; There is no team with an above .500 win percentage with less than 1000 passing yards and there are no teams with a win percentage under .400 with over 1000 passing yards.












